Random Generation of Rich \\ Abstract Data Type Values

### 2019-01-30 paper/synthesis.tex
General pass. Paper still needs compressing

In this work, we show how to naturally fit structural information beyond ADT
definitions into the prediction mechanism of branching processes.
%
It is our realization that it suffices to simply consider \emph{each different
  pattern matching and functional call as a kind of individual on its own}.
%
In that manner, we can extend our mean matrix $M$ by adding a row and a column
for each different pattern matching and function call as follows:

\newcommand{\ph}{\phantom{\square}}
\newcommand{\el}{\square}
\begin{equation*}
  \hspace{20pt}
  M = \ \
  \begin{tikzpicture}
    [ baseline=-0.65ex
    , every left delimiter/.style={xshift=.5em}
    , every right delimiter/.style={xshift=-.5em}
    ]
    \matrix
    [ matrix of math nodes
    , column sep=-1pt
    , row sep=-1pt
    , ampersand replacement=\&
    , left delimiter={[}
    , right delimiter={]}
    ] (m)
    {
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Text
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Sing
      \ph \& \ph \& \el \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Tag
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Join
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % simplify#1
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \el \& \ph \& \ph \\ % simplify#1
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % simplify#1
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % simplify#1
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % simplify#1
    };
    % Color squares
    \fill[red, opacity=0.075]
    (m-1-1.north west)
    -|| ([xshift=-1pt]m-1-3.north east)
    -|| ([xshift=-1pt, yshift=1pt]m-3-3.south east)
    -|| ([xshift=1pt, yshift=1pt]m-3-1.south west)
    -|| cycle;
    \fill[blue, opacity=0.075]
    ([yshift=-1pt]m-4-1.north west)
    -|| ([xshift=1pt]m-1-3.north east)
    -|| ([xshift=0pt, yshift=-1pt]m-1-9.south east)
    -|| ([xshift=0pt, yshift=0pt]m-9-9.south east)
    -|| ([xshift=-1pt, yshift=0pt]m-9-1.south east)
    -|| cycle;
    % Upper labels
    \node[above,text depth=1pt] at (m-1-1.north) {$\scriptstyle C_1$};
    \node[above,text depth=1pt] at (m-1-2.north) {$\scriptstyle \cdots$};
    \node[above,text depth=1pt] at (m-1-3.north) {$\scriptstyle C_i$};
    \node[above,text depth=1pt] at (m-1-4.north) {$\scriptstyle P_1$};
    \node[above,text depth=1pt] at (m-1-5.north) {$\scriptstyle \cdots$};
    \node[above,text depth=1pt] at (m-1-6.north) {$\scriptstyle P_j$};
    \node[above,text depth=1pt] at (m-1-7.north) {$\scriptstyle F_1$};
    \node[above,text depth=1pt] at (m-1-8.north) {$\scriptstyle \cdots$};
    \node[above,text depth=1pt] at (m-1-9.north) {$\scriptstyle F_k$};
    % Left labels
    \node[left,overlay] at (m-1-1.west) {$\scriptstyle C_1$};
    \node[left,overlay] at ([xshift=-4pt, yshift=2pt]m-2-1.west) {$\scriptstyle \vdots$};
    \node[left,overlay] at (m-3-1.west) {$\scriptstyle C_i$};
    \node[left,overlay] at (m-4-1.west) {$\scriptstyle P_1$};
    \node[left,overlay] at ([xshift=-4pt, yshift=2pt]m-5-1.west) {$\scriptstyle \vdots$};
    \node[left,overlay] at (m-6-1.west) {$\scriptstyle P_j$};
    \node[left,overlay] at (m-7-1.west) {$\scriptstyle F_1$};
    \node[left,overlay] at ([xshift=-4pt, yshift=2pt]m-8-1.west) {$\scriptstyle \vdots$};
    \node[left,overlay] at (m-9-1.west) {$\scriptstyle F_k$};
    % Dashed lines
    \draw[densely dashed] ([xshift=-2pt]m-3-1.south west) -- ([xshift=2pt]m-3-9.south east);
    \draw[densely dashed] ([xshift=-2pt]m-6-1.south west) -- ([xshift=2pt]m-6-9.south east);
    \draw[densely dashed] ([yshift=2pt]m-1-3.north east) -- ([yshift=-2pt]m-9-3.south east);
    \draw[densely dashed] ([yshift=2pt]m-1-6.north east) -- ([yshift=-2pt]m-9-6.south east);
    \draw[dotted, semithick] (m-3-1.west) -- ([xshift=3pt]m-3-3.west);
    \draw[dotted, semithick] (m-1-3.north) -- ([yshift=-3pt]m-3-3.north);
    \draw[dotted, semithick] (m-6-1.west) -- ([xshift=4pt]m-6-7.west);
    \draw[dotted, semithick] (m-1-7.north) -- ([yshift=-3pt]m-6-7.north);
  \end{tikzpicture}
\end{equation*}

Symbol $C_1 \cdots C_i$ denotes constructors, $P_1 \cdots P_j$ pattern
matchings, and $F_1 \cdots F_k$ function calls.
%
The red-light colored matrix is what we had before where the light-blue colored
cells are new---we recommend readers to get a colored copy of the paper.
%
% To extend $M$ to consider the new possible constructions, we need to calculate
% $M_{i,j}$ for every new combination of ``father'' and ``son'' constructions $i$
% and $j$, respectively.
% %

The new cells are filled as before: we need to consider the amount of holes when
generating partial pattern matching and function calls as well as their
individual probabilities.
%
% Fortunately, it is possible to follow the same pattern as we did on our original
% work, this time considering the amount of recursive pattern variables if it the
% case of a pattern matching value being a father individual, and the amount of
% recursive arguments if we deal with a abstract interface function call as a
% father individual.
%
For instance, if we consider $P_j$ as the second pattern of simplify and $F_1$
as function |div|, then the marked cell above has the value $2\cdot p_{div}$,
i.e., the amount of holes in a partially generated pattern |(Join (Join (Text s)
hole) hole)|, where |s| is some random string, times the probability to generate
a call to function |div|.
%
The rest follows analogously.

@@

As another contribution, we found that the whole prediction process can be
factored in terms of two vectors $\beta$ and $\mathcal{P}$, such that $\beta$
represents the number of recursive sub-terms of each construction that we
generate, whereas $\mathcal{P}$ simply represents the probability of generating
that partial ADT value.%  (this probability can be easily calculated based on the
% generation frequencies used in the random generator).
%
Then, the equation \ref{eqn:prediction} can be rewriten as:
%
\begin{align*}
  E[G_n]^T = \beta^T \cdot (\beta \cdot \mathcal{P})^{n}
\end{align*}
%
The vector $\beta$ can be obtained by analyzing the shape (namely holes) of each
source of structural information that we consider, and joining them together.
%
For instance, $\beta$ and $\mathcal{P}$ for our generation specification of HTML
values are as follows (where dashed lines indicate the composition of the
diffent sub-vectors that can be calculated separately):

@@

Note that by varying the shape of the vector $\mathcal{P}$ we can tune the
distribution of our random genrator in a way that can be always characterized
and predicted.
%
\dragenp follows a similar approach as \dragen and uses an heuristic to tune the
generation probabilities of each source of structural information.
%
This is done by runing a simulation-based optimization process at compile-time.
%
This process is parametrized by the desired distribution of values set by the
user.
%
In this manner, developers can specify, for instance, a uniform distribution of
data constructors, pattern matching values and function calls or, alternative, a
distribution of values with some constructions appearing in a different
proportion as others, e.g., two times more functions calls to |div| than |Join|
constructors.

>>>>

In this work, we show how to naturally fit structural information beyond ADT
definitions into the prediction mechanism of branching processes.
%
It is our realization that it suffices to simply consider \emph{each different
  pattern matching and function call as a kind of individual on its own}.
%
In that manner, we can extend our mean matrix $M$ by adding a row and a column
for each different pattern matching and function call as follows:
%
\newcommand{\ph}{\phantom{\square}}
\newcommand{\el}{\square}
\begin{equation*}
  \hspace{20pt}
  M = \ \
  \begin{tikzpicture}
    [ baseline=-0.65ex
    , every left delimiter/.style={xshift=.5em}
    , every right delimiter/.style={xshift=-.5em}
    ]
    \matrix
    [ matrix of math nodes
    , column sep=-1pt
    , row sep=-1pt
    , ampersand replacement=\&
    , left delimiter={[}
    , right delimiter={]}
    ] (m)
    {
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Text
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Sing
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Tag
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Join
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % simplify#1
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \el \& \ph \& \ph \\ % simplify#1
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % simplify#1
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % simplify#1
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % simplify#1
    };
    % Color squares
    \fill[red, opacity=0.075]
    (m-1-1.north west)
    -|| ([xshift=-1pt]m-1-3.north east)
    -|| ([xshift=-1pt, yshift=1pt]m-3-3.south east)
    -|| ([xshift=1pt, yshift=1pt]m-3-1.south west)
    -|| cycle;
    \fill[blue, opacity=0.075]
    ([yshift=-1pt]m-4-1.north west)
    -|| ([xshift=1pt]m-1-3.north east)
    -|| ([xshift=0pt, yshift=-1pt]m-1-9.south east)
    -|| ([xshift=0pt, yshift=0pt]m-9-9.south east)
    -|| ([xshift=-1pt, yshift=0pt]m-9-1.south east)
    -|| cycle;
    % Upper labels
    \node[above,text depth=1pt] at (m-1-1.north) {$\scriptstyle C_1$};
    \node[above,text depth=1pt] at (m-1-2.north) {$\scriptstyle \cdots$};
    \node[above,text depth=1pt] at (m-1-3.north) {$\scriptstyle C_i$};
    \node[above,text depth=1pt] at (m-1-4.north) {$\scriptstyle P_1$};
    \node[above,text depth=1pt] at (m-1-5.north) {$\scriptstyle \cdots$};
    \node[above,text depth=1pt] at (m-1-6.north) {$\scriptstyle P_j$};
    \node[above,text depth=1pt] at (m-1-7.north) {$\scriptstyle F_1$};
    \node[above,text depth=1pt] at (m-1-8.north) {$\scriptstyle \cdots$};
    \node[above,text depth=1pt] at (m-1-9.north) {$\scriptstyle F_k$};
    % Left labels
    \node[left,overlay] at (m-1-1.west) {$\scriptstyle C_1$};
    \node[left,overlay] at ([xshift=-4pt, yshift=2pt]m-2-1.west) {$\scriptstyle \vdots$};
    \node[left,overlay] at (m-3-1.west) {$\scriptstyle C_i$};
    \node[left,overlay] at (m-4-1.west) {$\scriptstyle P_1$};
    \node[left,overlay] at ([xshift=-4pt, yshift=2pt]m-5-1.west) {$\scriptstyle \vdots$};
    \node[left,overlay] at (m-6-1.west) {$\scriptstyle P_j$};
    \node[left,overlay] at (m-7-1.west) {$\scriptstyle F_1$};
    \node[left,overlay] at ([xshift=-4pt, yshift=2pt]m-8-1.west) {$\scriptstyle \vdots$};
    \node[left,overlay] at (m-9-1.west) {$\scriptstyle F_k$};
    % Dashed lines
    \draw[densely dashed] ([xshift=-2pt]m-3-1.south west) -- ([xshift=2pt]m-3-9.south east);
    \draw[densely dashed] ([xshift=-2pt]m-6-1.south west) -- ([xshift=2pt]m-6-9.south east);
    \draw[densely dashed] ([yshift=2pt]m-1-3.north east) -- ([yshift=-2pt]m-9-3.south east);
    \draw[densely dashed] ([yshift=2pt]m-1-6.north east) -- ([yshift=-2pt]m-9-6.south east);
    % \draw[dotted, semithick] (m-3-1.west) -- ([xshift=3pt]m-3-3.west);
    % \draw[dotted, semithick] (m-1-3.north) -- ([yshift=-3pt]m-3-3.north);
    \draw[dotted, semithick] (m-6-1.west) -- ([xshift=4pt]m-6-7.west);
    \draw[dotted, semithick] (m-1-7.north) -- ([yshift=-3pt]m-6-7.north);
  \end{tikzpicture}
\end{equation*}

Symbol $C_1 \cdots C_i$ denotes constructors, $P_1 \cdots P_j$ pattern
matchings, and $F_1 \cdots F_k$ function calls.
%
The red-light colored matrix is what we had before where the light-blue colored
cells are new---we encourage readers to obtain a colored copy of this work.
%
% To extend $M$ to consider the new possible constructions, we need to calculate
% $M_{i,j}$ for every new combination of ``father'' and ``son'' constructions $i$
% and $j$, respectively.
% %

The new cells are filled as before: we need to consider the amount of holes when
generating partial pattern matching and function calls as well as their
individual probabilities.
%
% Fortunately, it is possible to follow the same pattern as we did on our original
% work, this time considering the amount of recursive pattern variables if it the
% case of a pattern matching value being a father individual, and the amount of
% recursive arguments if we deal with a abstract interface function call as a
% father individual.
%
For instance, if we consider $P_j$ as the second pattern of simplify and $F_1$
as function |div|, then the marked cell above has the value $2\cdot p_{div}$,
i.e., the amount of holes in a partially generated pattern |(Join (Join (Text s)
hole) hole)|, where |s| is some random string, times the probability to generate
a call to function |div|.
%
The rest of this matrix can be computed analogously.

@@

As another contribution, we found that the whole prediction process can be
factored in terms of two vectors $\beta$ and $\mathcal{P}$, such that $\beta$
represents the number of recursive sub-terms of each construction that we
generate (i.e., data constructors, pattern matching values or function calls),
whereas $\mathcal{P}$ simply represents the probability of generating that
partial ADT value.% (this probability can be easily calculated based on the
% generation frequencies used in the random generator).
%
Then, the equation \ref{eqn:prediction} can be rewriten as:
%
\begin{align*}
  E[G_n]^T = \beta^T \cdot (\beta \cdot \mathcal{P})^{n}
\end{align*}
%
The vector $\beta$ can be obtained by analyzing the shape (namely holes) of each
source of structural information that we consider, and joining them together.
%
For instance, $\beta$ and $\mathcal{P}$ for our generation specification of HTML
values are as follows (where dashed lines indicate the composition of the
diffent sub-vectors that are calculated separately):

@@

Note that by varying the shape of the vector $\mathcal{P}$ we can tune the
distribution of our random genrator in a way that can be always characterized
and predicted.
%
\dragenp follows a similar approach as \dragen and uses an heuristic to tune the
generation probabilities of each source of structural information.
%
This is done by runing a simulation-based optimization process at compile-time.
%
This process is parametrized by the desired distribution of values set by the
user.
%
In this manner, developers can specify, for instance, a uniform distribution of
data constructors, pattern matching values and function calls or, alternatively,
a distribution of values with some constructions appearing in a different
proportion as others, e.g., two times more functions calls to |div| than |Join|
constructors.
